---
title: "Covid91 vaccine study Final2020"
author: "Kathleen Giba"
date: "27, October 2020"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode aids to say covid and then remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine.

$H_0$: The Null Hypothesis is that among the target population, the vaccine has no effect on whether one contracts the virus or not.  

$H_A$: The alternative Hypothesis, is that among the target population, the vaccine has an effect on whether one contracts the virus or not.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->  

Every question is of the Cat~Cat type.  

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

#Graphical Descriptive Results for Gals.  

```{r}
DataF = filter(FinalData,sex=="gal")
barchartGC(~infected + treatment,data=DataF)
barchartGC(~infected + treatment,data=DataF, type="percent",main="gals")

``` 

Among the females who contracted Covid, twice as many took the placebo. Therefore it appears that the vaccine does have a positive effect.  

## Numerical Summeries for Gals.  

```{r}
table1 <- xtabs(~infected + treatment,data=DataF)
rowPerc(table1)
colPerc(table1)
```

You can see from the numerical summaries that the the amount of infected 33% took the drug and 66% took the placebo. The drug works for the gals.   

## Inferential for Gals.  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The fisher test shows us that the odds of 0.4889403 % of gal who took the placebo got covid compared to taking the drug.  

```{r}
DataG = filter(FinalData,sex=="guy")
barchartGC(~infected + treatment,data=DataG)
barchartGC(~infected + treatment,data=DataG, type="percent",main="guys")

```

It appears that the vaccine has a negative effect on the guys, It seems that they are more likely to get Covid from the vaccine.  

## Numerical Summeries for Guys.  

```{r}
table1 <- xtabs(~infected + treatment,data=DataG)
rowPerc(table1)
colPerc(table1)
```

As you can see the amount of infected is 58.16% who took the drug and 41.84 % took the placebo. therefor taking the drug for guys has a negative effect.  

## Inferential for Guys.

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The odds of 1.473872 more likely for guys who took the drug to get covid than taking the placebo.    

```{r}
DataLGBTQ = filter(FinalData,LGBTQ=="gay")
barchartGC(~infected + treatment,data=DataLGBTQ)
barchartGC(~infected + treatment,data=DataLGBTQ, type="percent",main="gays")

```
  
It appears that the vaccine has a negative effect on the LGBTQ, It seems that they are more likely to get Covid from the vaccine.    

## Numerical Summeries for Gays.

```{r}
DataDrugUser = filter(FinalData,DrugUser=="yes")
barchartGC(~infected + treatment,data=DataDrugUser)
barchartGC(~infected + treatment,data=DataDrugUser, type="percent",main="drugs")

```
  
Among the Drug Users who contracted Covid, twice as many took the placebo. Therefore it appears that the vaccine does have a positive effect.       
 
  
# Discussion and Conclusion



<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
